---
title: "R Notebook"
output: html_notebook
---

# Loading libraries
```{r}
require(revgeo)
```

# Load researcher trajectory data
```{r}
t <- read.csv('trajectory_oh_1_person_lat_lon.csv')
```

```{r}
head(t)
```

Some statistics on the daily travelled distance

```{r}
hist.0(t$distance, 50, xlab='Distance travelled')
```

# Number of days spent in each location

We look at the total number of days spent at a same lat,lon coordinates. Ranking by most visited to least visited places, we look at the distribution of number of days to sort out "home" and "family" locations from pure travels.
```{r}
sp <- split(t, paste(t$latitude, t$longitude))
Ndays <- unlist(lapply(sp, nrow))
inds <- order(-Ndays)
plot.0(Ndays[inds],
       xlab='Location ranked by number of days',
       ylab='Number of days',
       log='xy'
       )
```

# Heatmap

Here we use a heatmap to represent in time (columns) which places are visited (rows). Red denotes visited, white not visited.

```{r}
locs_a <- paste(t$latitude, t$longitude)
locs <- unique(paste(t$latitude, t$longitude))
M <- NULL
for (loc in locs){
  M <- rbind(M, as.numeric(locs_a == loc))
}
colnames(M) <- t$date
```

```{r}
pdf('plotHeatmap.pdf')
heatmap.0(M, cexRow = 0.6, cexCol = 0.2, xlab='Time')
dev.off()
```

# Reverse geolocation

We use the revgeo library to get full addresses from lat,lon data using Open Street Maps Photon API. We use unique lat,lon coordinates to avoid overflooding the API.
```{r, include=FALSE}
locs <- unique(paste(t$latitude, t$longitude))
locs_sp <- t(sapply(locs, function(x) strsplit(x,' ')[[1]]))
geos <- revgeo(longitude=locs_sp[,2], latitude=locs_sp[,1], output='hash')
```

```{r}
save(geos, file='geolocations.RData')
```

# Saving full dataframe

We save the previous information to the initial dataframe for further manipulation
```{r}
t$city <- rep(NA, nrow(t))
t$state <- rep(NA, nrow(t))
t$zip <- rep(NA, nrow(t))
t$country <- rep(NA, nrow(t))
for (i in 1:length(locs)){
  inds <- which(locs_a == locs[i])
  t$city[inds] <- geos$city[i]
  t$state[inds] <- geos$state[i]
  t$zip[inds] <- geos$zip[i]
  t$country[inds] <- geos$country[i]
}

```

```{r}
write.csv(t, file='trajectory_oh_1_person_lat_lon_with_geo.csv')
```

# Repeat analysis for city, zip, state levels

```{r}
sp <- split(t, t$zip)
Ndays <- unlist(lapply(sp, nrow))
inds <- order(-Ndays)
plot.0(Ndays[inds],
       xlab='ZIP code ranked by number of days',
       ylab='Number of days',
       log='xy'
       )
```

```{r}
locs_a <- t$zip
locs <- unique(locs_a)
M <- NULL
for (loc in locs){
  M <- rbind(M, as.numeric(locs_a == loc))
}
colnames(M) <- t$date
rownames(M) <- locs

pdf('plotHeatmap_zip.pdf')
heatmap.0(M, cexRow = 0.6, cexCol = 0.2, xlab='Time')
dev.off()
```


```{r}
sp <- split(t, t$city)
Ndays <- unlist(lapply(sp, nrow))
inds <- order(-Ndays)
plot.0(Ndays[inds],
       xlab='City ranked by number of days',
       ylab='Number of days',
       log='xy'
       )
```

```{r}
locs_a <- t$city
locs <- unique(locs_a)
M <- NULL
for (loc in locs){
  M <- rbind(M, as.numeric(locs_a == loc))
}
colnames(M) <- t$date
rownames(M) <- locs

pdf('plotHeatmap_city.pdf')
heatmap.0(M, cexRow = 0.6, cexCol = 0.2, xlab='Time')
dev.off()
```


```{r}
sp <- split(t, t$state)
Ndays <- unlist(lapply(sp, nrow))
inds <- order(-Ndays)
plot.0(Ndays[inds],
       xlab='State ranked by number of days',
       ylab='Number of days',
       log='xy'
       )
```


```{r}
locs_a <- t$state
locs <- unique(locs_a)
M <- NULL
for (loc in locs){
  M <- rbind(M, as.numeric(locs_a == loc))
}
colnames(M) <- t$date
rownames(M) <- locs

pdf('plotHeatmap_state.pdf')
heatmap.0(M, cexRow = 0.6, cexCol = 0.2, xlab='Time')
dev.off()
```



```{r}
sp <- split(t, t$country)
Ndays <- unlist(lapply(sp, nrow))
inds <- order(-Ndays)
plot.0(Ndays[inds],
       xlab='Country ranked by number of days',
       ylab='Number of days',
       log='xy'
       )
```



# Splitting data by location

We split by city

```{r}
sp <- split(t$distance, t$city)
mu <- unlist(lapply(sp, mean))
se <- unlist(lapply(sp, se.0))
inds <- order(-mu)

Ndays <- unlist(lapply(sp[inds], length))
cols <- rep('gray', length(Ndays))
cols[order(-Ndays)[1:3]] <- 'red'


```

```{r}
boxplot.0(sp[inds], 
          mar=c(5,5,5,5), 
          las=3, 
          cex.axis = 0.6, 
          ylab='Distance travelled',
          main='Cities',
          col=cols
          )
```

